---
title: "R Notebook"
output: html_notebook
---
### load packages

```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
```


### additional packages being used
```{r message=FALSE, warning=FALSE}
library(tmap)
library(sf)
library(tidycensus)
```

### load testing data

```{r message=FALSE, warning=FALSE}
df1=readxl::read_excel(here("data","raw","tot_tests_neighborhood.xlsx")) %>% 
  clean_names() %>%
  #mutate(date_reported= as_datetime(date_reported))
  separate(date_reported,into=c("date","hr"),sep = " ")%>%
  mutate(date_report=as_date(date))%>%
  group_by(neighborhood)%>%
  mutate(across(total_tests, ~ .-c(0,lag(.)[-1])))
```


### Load Mapping Data

```{r message=FALSE, warning=FALSE}
neigh=st_read(here("data","raw","dc_neigh.geojson")) %>% clean_names()
class(neigh)
```

### celect and clean data

```{r}
df_tests=df1 %>%
  filter(as_date(date_report) == "2021/10/27") %>% 
  separate(neighborhood,into=c("code","name"),sep = ":") %>% 
  mutate(code=case_when(
    code=="N35" ~"N0",
    TRUE ~ code
  ))
```

### join mapping data with test data

```{r message=FALSE, warning=FALSE}
neigh2=left_join(neigh,df_tests,by=c("code")) 
tmap_mode("view")
tm_shape(neigh2) +tm_polygons("total_tests",alpha=.5)
```

### load positive cases data
```{r message=FALSE, warning=FALSE}
df_c=readxl::read_excel(here("data","raw","neigh_cases.xlsx"),
                            col_types = c("date", "text", "numeric")) %>% 
  clean_names() 
df_cases=df_c %>%
  filter(as_date(date) == "2021/10/27") %>% 
  separate(neighborhood,into=c("code","name"),sep = ":") %>%
  mutate(code=case_when(
    code=="N35" ~"N0",
    TRUE ~ code
  ))
```

### Join covid data

```{r}
neigh3=left_join(neigh2,df_cases,by=c("code")) 
tmap_mode("view")
tm_shape(neigh3) +tm_polygons("total_positives",alpha=.5)
```

### graph data

```{r}
 tm_shape(neigh3)+tm_polygons(c("total_positives","total_tests"),alpha=.5)
```

### get census Data
```{r}
census_api_key("2c8b9d5c4902b7efb4e1f98b2c23692cb1b73e95")

v20 = load_variables(2018,"acs5")
df_cencus=get_acs(geography = "tract",
                  variables=c("pop"="B01001_001"),
                  state="DC",geometry=TRUE,year=2018) 

df_cens=df_cencus %>% select(-moe) %>% spread(variable,estimate) 
df_cens_adj=df_cens %>% st_transform(4326)
```

### Join Census data to mapping data
```{r}
df_j=st_join(df_cens_adj,neigh3,largest=TRUE)
```

### Summarize population for each neighbourhood
```{r}
df3=df_j %>% select(pop,code) %>%
  group_by(code) %>%
  summarise(pop_n=sum(pop))
            
            
```

### Join population data and generate results

Removed any extreme outliers and generated map to compare testing rates and covid rates
```{r}
df4=left_join(neigh3,df3 %>% st_set_geometry(NULL))

df4=df4 %>% mutate( covid_rate=total_positives/pop_n, test_rate=total_tests/pop_n)
df4 %>% filter(!(code %in% c("N0","N15","N24"))) %>% tm_shape()+tm_polygons(c("covid_rate","test_rate"))
```



