---
title: "Comparing Covid Tests and Covid Positives using mapping data"
output: html_notebook
---

Through mapping both covid tests and positives adjusted for population, we hope to fine if any relation between the two exists. It may be possible that more tests mean more covid.

### Load packages

The following packages are used:

knitr-- load datasets

tidverse-- mutate data

janitor-- clean names

lubridate -- work with dates

here-- give file locations

```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
```


### Additional packages being used

tmap-- for mapping purposes

sf-- joining spacial data and generating plots

tidycensus-- loading census data

```{r message=FALSE, warning=FALSE}
library(tmap)
library(sf)
library(tidycensus)
```

### Load testing data

The test data is loaded into R, and mutated to become non-cumulative

```{r message=FALSE, warning=FALSE}
df1=readxl::read_excel(here("data","raw","tot_tests_neighborhood.xlsx")) %>% 
  clean_names() %>%
  #mutate(date_reported= as_datetime(date_reported))
  separate(date_reported,into=c("date","hr"),sep = " ")%>%
  mutate(date_report=as_date(date))%>%
  group_by(neighborhood)%>%
  mutate(across(total_tests, ~ .-c(0,lag(.)[-1])))
```


### Load Mapping Data

Loads in the Health neighbourhood geojson file.

```{r message=FALSE, warning=FALSE}
neigh=st_read(here("data","raw","dc_neigh.geojson")) %>% clean_names()
class(neigh)
```

### Select and clean data

Cleans testing data so it can be joined with the health neighborhood spacial data

```{r}
df_tests=df1 %>%
  filter(as_date(date_report) == "2021/10/27") %>% 
  separate(neighborhood,into=c("code","name"),sep = ":") %>% 
  mutate(code=case_when(
    code=="N35" ~"N0",
    TRUE ~ code
  ))
```

### Join mapping data with test data

Joins data and generates plot

```{r message=FALSE, warning=FALSE}
neigh2=left_join(neigh,df_tests,by=c("code")) 
tmap_mode("view")
tm_shape(neigh2) +tm_polygons("total_tests",alpha=.5)
```

### Load positive cases data

Reads in positive coivd cases data and cleans it so it can be joined to the above datasets

```{r message=FALSE, warning=FALSE}
df_c=readxl::read_excel(here("data","raw","neigh_cases.xlsx"),
                            col_types = c("date", "text", "numeric")) %>% 
  clean_names() 
df_cases=df_c %>%
  filter(as_date(date) == "2021/10/27") %>% 
  separate(neighborhood,into=c("code","name"),sep = ":") %>%
  mutate(code=case_when(
    code=="N35" ~"N0",
    TRUE ~ code
  ))
```

### Join covid data

```{r}
neigh3=left_join(neigh2,df_cases,by=c("code")) 
tmap_mode("view")
tm_shape(neigh3) +tm_polygons("total_positives",alpha=.5)
```

### Graph data

```{r}
 tm_shape(neigh3)+tm_polygons(c("total_positives","total_tests"),alpha=.5)
```

### Get census Data

Census data is used to normalize both tests and COVID data for population

```{r}
census_api_key("2c8b9d5c4902b7efb4e1f98b2c23692cb1b73e95")

v20 = load_variables(2018,"acs5")
df_cencus=get_acs(geography = "tract",
                  variables=c("pop"="B01001_001"),
                  state="DC",geometry=TRUE,year=2018) 

df_cens=df_cencus %>% select(-moe) %>% spread(variable,estimate) 
df_cens_adj=df_cens %>% st_transform(4326)
```

### Join Census data to mapping data
```{r}
df_j=st_join(df_cens_adj,neigh3,largest=TRUE)
```

### Summarize population for each neighbourhood
```{r}
df3=df_j %>% select(pop,code) %>%
  group_by(code) %>%
  summarise(pop_n=sum(pop))
            
            
```

### Join population data and generate results

Removed any extreme outliers and generated map to compare testing rates and covid rates

```{r}
df4=left_join(neigh3,df3 %>% st_set_geometry(NULL))

df4=df4 %>% mutate( covid_rate=total_positives/pop_n, test_rate=total_tests/pop_n)
df4 %>% filter(!(code %in% c("N0","N15","N24"))) %>% tm_shape()+tm_polygons(c("covid_rate","test_rate"))
```


