---
title: "Bike Analysis"
author: "Shreejit Poudyal"
date: "11/17/2021"
output: html_document
---

### Loading Packages

The following packages are used: 

tidyverse: mutate data
readr: Read csv, xlxs from local computer or web
lubridate: manipulate dates
janitor: clean names
ggplot2: create visualizations
sf: load kml files
tmap: mapping tools
tidycensus:extract available census data

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(lubridate)
library(janitor)
library(ggplot2)
library(here)
library(sf) # working with simple features - geospatial
library(tmap)
library(tidycensus)
```

### Data

The data was received from https://opendata.dc.gov/search?collection=Dataset&q=crime, where we extracted data for the years 2017. 2018, 2019, 2020 and 2021 (till November). The data is free to use for anyone willing to analyse it for their own purposes.

```{r, include = FALSE}
data <- read.csv("https://raw.githubusercontent.com/jskufca/DS241_bikes_2021/shreejit/data/crime_data/crime_processed/final.csv")
```

```{r}
census_geo = st_read(here("data","crime_data", "crime_raw", "Census_Tracts_in_2020.geojson")) %>% clean_names()
crime = st_read(here("data","crime_data", "crime_raw", "crime_2018.geojson")) %>% clean_names()
class(crime)
plot(crime)
plot(census_geo)
```

### Prepping Data

I renamed various columns so that it is easier for me to know and understand the variables to better use it for visualizations. One of which includes using the `as_datettime` function that converts date format input to date-time format (`YYYY-MM-DD HH:MM:SS`). I renames various variables and formatted the data to something suitable and easily understandable.

```{r, include = FALSE}
main <- data %>%
  select(CCN:OBJECTID) %>%
  rename(complaint_num = CCN) %>%
  rename(report_date = REPORT_DAT) %>%
  rename(shift = SHIFT) %>%
  rename(method = METHOD) %>%
  rename(offense = OFFENSE) %>%
  rename(block = BLOCK) %>%
  rename(x_block = XBLOCK) %>%
  rename(y_block = YBLOCK) %>%
  rename(ward = WARD) %>%
  rename(anc = ANC) %>% # Advisory Neighborhood Commission
  rename(district = DISTRICT) %>%
  rename(psa = PSA) %>% # Police Service Area
  rename(nbhd_cluster = NEIGHBORHOOD_CLUSTER) %>%
  rename(block_grp = BLOCK_GROUP) %>%
  rename(census_tract = CENSUS_TRACT) %>%
  rename(precinct = VOTING_PRECINCT) %>%
  rename(latitude = LATITUDE) %>%
  rename(longitude = LONGITUDE) %>%
  rename(bid = BID) %>% # Business Improvement District
  rename(start = START_DATE) %>%
  rename(end = END_DATE) %>%
  rename(object_id = OBJECTID) %>%
  mutate(report_date = substr(report_date, 1, nchar(report_date)-3),
         report_date = as_datetime(report_date),
         start = substr(start, 1, nchar(start)-3),
         start = as_datetime(start),
         end = substr(end, 1, nchar(end)-3),
         end = as_datetime(end),
         nbhd_cluster = as.numeric(substring(nbhd_cluster, 9)),
         precinct = as.numeric(substring(precinct, 10)),
         block_grp = as.numeric(substring(block_grp, 8)),
         census_tract = as.numeric(census_tract),
         shift=str_to_sentence(shift),
         method=str_to_sentence(method),
         offense=str_to_sentence(offense),
         bid=str_to_title(bid))
```

### Making Visualizations form the Data

#### Time of the day

```{r}
ggplot(main, aes(fct_infreq(shift))) +
  geom_bar() +
  labs(
    x="Time of Day",
    y="Number of Crimes Reported",
    title="Cases count by time of day (since 2017)")+
  theme(legend.position = "none")
```

#### Crimes by Hour

The following `ggplot` code shows us a agraph with total crimes happened in each hour since 2017.

```{r}
ggplot(main, aes(hour(report_date)), fill=shift) +
  geom_bar() +
  labs(
    x="Hour of the Day",
    y="Number of Crimes Reported",
    title="Crimes by Hour (since 2017)")+
  theme(legend.position = "none")
```

#### Method

```{r}
ggplot(main, aes(x=fct_infreq(method), fill=method)) +
  geom_bar() +
  labs(
    x="Weapon of use",
    y="Number of Crimes Reported",
    fill="Method",
    title="Method of Crime") +
  scale_y_continuous(labels = scales::comma)+
  theme(legend.position = "none")
```

#### Offense

```{r}
ggplot(main, aes(x=fct_infreq(offense), fill=offense)) +
  geom_bar(stat='count') +
  #coord_cartesian(ylim=c(0,100)) + # Number of Arson crimes   is very small (<50)
  labs(
    x="Crime",
    y="Number of Crimes Reported",
    fill="Offense",
    title="Criminal Offense") +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) +
    theme(legend.position = "none")
```

#### Total Crimes since 2017 (by month)

```{r}
ggplot(main, aes(factor(month(report_date, label=TRUE)))) +
  geom_bar() +
  labs(
    x="Year",
    y="Number of Crimes Reported") +
    theme(legend.position = "none")
```

#### Yearwise Number of crimes

```{r}
ggplot(main, aes(factor(year(report_date)))) +
  geom_bar() +
  labs(
    x="Year",
    y="Number of Crimes Reported") +
    theme(legend.position = "none")
```
#### Did COVID impact crimes in December 2019 ?

```{r}
data_2019 <- main %>%
  filter(year(report_date)==2019)

ggplot(data_2019, aes(factor(month(report_date, label=TRUE)))) +
  geom_bar() +
  labs(
    x="Year",
    y="Number of Crimes Reported",
    title="Number of Crimes Reported per month in 2019") +
    theme(legend.position = "none")
```
#### How about 2020 ?

```{r}
data_2020 <- main %>%
  filter(year(report_date)==2020)

ggplot(data_2020, aes(factor(month(report_date, label=TRUE)))) +
  geom_bar() +
  labs(
    x="Year",
    y="Number of Crimes Reported",
    title="Number of Crimes Reported per month in 2020") +
    theme(legend.position = "none")
```

#### Data for 2021

```{r}
data_2021 <- main %>%
  filter(year(report_date)==2021)

ggplot(data_2021, aes(factor(month(report_date, label=TRUE)))) +
  geom_bar() +
  labs(
    x="Year",
    y="Number of Crimes Reported",
    title="Number of Crimes Reported per month in 2020") +
    theme(legend.position = "none")
```     