---
title: "Team 1 Mapping"
author: "Jackie, Mark, Jackson"
date: "11/5/2021"
output: html_notebook
---

##Load Libraries

Here we loaded a few libraries we will be needing
```{r eval = FALSE, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(lubridate)
library(janitor)
library(readr)
library(sf)
library(tmap)
```


##Our Goals
Team 1 - Mapping

Task 1 - Revisit what we have done with maps.   Identify where we have made maps in the class, and what packages we have used. 

Task 2 - Make a plot of at least some subset of the datapoints from the bike share data.

We will want to use the  `sf` package.   See if you can find a tutorial that is "our speed."

We will want to use the `tmap` package.   See if you can find a tutorial that is "our speed."

##Answering Task 1

-We have made maps in class. 

-We made a map in Lab 04. There were maps per state, and maps of the entire US.

-Some classmates were able to generate a plot in Lab 04 on an actual interactive map. This map included borders, and ability to zoom in and out. (Was it
Shreejit?)
How was this done? It could be a useful tool.

-The packages used in lab 04 were tidyverse and and dsbox

##Task 2
Here the goal is to make a plot of a subset of the datapoints from the bike share data.

Our subset will be the locations of pick-ups and drop-offs for Saturday September 4, 2021

```{r}
df1a <-  unzip(202108-capitalbikeshare-tripdata.zip) %>%
  read_csv("https://s3.amazonaws.com/capitalbikeshare-data/index.html.csv") %>% clean_names
```


```{r download_data}
df1 <- read_csv("U:/Data Science 241/202109-capitalbikeshare-tripdata.csv") %>% clean_names
```

Here We want to get the data filtered to just Sep 4
```{r manipulate_data}
df1a = df1 %>% 
  select(rideable_type, start = started_at, end = ended_at, start_station_name:end_lng) %>%
  pivot_longer(cols =  start:end,  names_to = "type", values_to = "time")%>%
  arrange(time) %>%
  mutate(inc = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  ), riders = cumsum(inc)) %>%
  mutate(day_of_year = yday(time)) %>%
  mutate(day_since_sep1 = day_of_year-243) %>%
  filter(day_since_sep1 == 4)
```

Here is the visualization
```{r visualization}
ggplot(df1a, mapping = aes(x = start_lat, y = start_lng)) +
  geom_point()

ggplot(df1a, mapping = aes(x = end_lat, y = end_lng)) +
  geom_point()
```
It would be beneficial if this plot were overlaid over some map of the city. Right now the points do not mean much.

##SF File Work

```{r}
sf_neighborhoods_df <- sf::read_sf("U:/Data Science 241/DS241_bikes_2021/DC_Health_Planning_Neighborhoods")
 View(sf_neighborhoods_df)
```

```{r}
plot(sf_neighborhoods_df) 
```
