---
title: "Brennas file"
output: html_notebook
---
https://github.com/jskufca/DS241_bikes_2021

Taken from https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html

##Hello world
```{r}
library(tmap)
data("World")

tm_shape(World) +
    tm_polygons("HPI")
```

The object World is a spatial object of class sf from the sf package; it is a data.frame with a special column that contains a geometry for each row, in this case polygons. In order to plot it in tmap, you first need to specify it with tm_shape. Layers can be added with the + operator, in this case tm_polygons. There are many layer functions in tmap, which can easily be found in the documentation by their tm_ prefix. See also ?'tmap-element'.


##Interactive Maps
Each map can be plotted as a static image or viewed interactively using "plot" and "view" modes, respectively. The mode can be set with the function tmap_mode, and toggling between the modes can be done with the ‘switch’ ttm() (which stands for toggle thematic map.
```{r}
tmap_mode("view")

tm_shape(World) +
    tm_polygons("HPI")
```

##Multiple shapes and layers
A shape is a spatial object (with a class from sf, sp, stars, or raster). Multiple shapes and also multiple layers per shape can be plotted:
```{r}
data(World, metro, rivers, land)

tmap_mode("plot")
## tmap mode set to plotting
tm_shape(land) +
    tm_raster("elevation", palette = terrain.colors(10)) +
tm_shape(World) +
    tm_borders("white", lwd = .5) +
    tm_text("iso_a3", size = "AREA") +
tm_shape(metro) +
    tm_symbols(col = "red", size = "pop2020", scale = .5) +
tm_legend(show = FALSE)
```

## Facets
Facets can be created in three ways:

1. By assigning multiple variable names to one aesthetic (in this example the first argument of tm_polygons:
```{r}
tmap_mode("view")
tm_shape(World) +
    tm_polygons(c("HPI", "economy")) +
    tm_facets(sync = TRUE, ncol = 2)
```

2. By splitting the spatial data with the by argument of tm_facets:
```{r}
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)

NLD_muni$perc_men <- NLD_muni$pop_men / NLD_muni$population * 100

tm_shape(NLD_muni) +
    tm_polygons("perc_men", palette = "RdYlBu") +
    tm_facets(by = "province")
```

3. By using the tmap_arrange function:
```{r}
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)
tm1 <- tm_shape(NLD_muni) + tm_polygons("population", convert2density = TRUE)
tm2 <- tm_shape(NLD_muni) + tm_bubbles(size = "population")

tmap_arrange(tm1, tm2)
## Legend labels were too wide. Therefore, legend.text.size has been set to 0.27. Increase legend.width (argument of tm_layout) to make the legend wider and therefore the labels larger.
## The legend is too narrow to place all symbol sizes.
```
##Basemaps and overlay tile maps
Tiled basemaps can be added with the layer function tm_basemap. Semi-transparent overlay maps (for example annotation labels) can be added with tm_tiles.
```{r}
tmap_mode("view")
tm_basemap("Stamen.Watercolor") +
tm_shape(metro) + tm_bubbles(size = "pop2020", col = "red") +
tm_tiles("Stamen.TonerLabels")
```

##Options and styles
The functions tm_layout and tm_view are used to specify the map layout and the interactive aspects respectively. These functions can be used in the same way as the layer functions, e.g.
```{r}
tmap_mode("plot")
## tmap mode set to plotting

tm_shape(World) +
    tm_polygons("HPI") +
tm_layout(bg.color = "skyblue", inner.margins = c(0, .02, .02, .02))
```

These options, as well as a couple of others, can also be set within with tmap_options, which works in the same way as the base R function options. The main advantage is that these options are set globally, so they do not have to be specified in each map, for the duration of the session

```{r}
tmap_options(bg.color = "black", legend.text.color = "white")

tm_shape(World) +
    tm_polygons("HPI", legend.title = "Happy Planet Index")
```

A style is a certain configuration of the tmap options

```{r}

tmap_style("classic")
## tmap style set to "classic"
## other available styles are: "white", "gray", "natural", "cobalt", "col_blind", "albatross", "beaver", "bw", "watercolor"

tm_shape(World) +
    tm_polygons("HPI", legend.title = "Happy Planet Index")

```
```{r}
# see what options have been changed
tmap_options_diff()

# reset the options to the default values
tmap_options_reset()

```

