<<<<<<< HEAD
---
title: "dc_covid"
author: "Jackie Budka"
date: "11/29/2021"
output: html_notebook
---



```{r load packages}
library(tidyverse)
library(tidymodels)
library(dsbox)
library(httr)
library(jsonlite)
library(lubridate)
library(sf)
library(tmap)
library(ggplot2)
```

```{r}
# url <- "https://opendata.arcgis.com"
# path <- "datasets/62f6f28dbfc74ae489fa3dcda9e94744_28.geojson"
# raw_data <- GET(url,path)
```
```{r}
neighborhoods_shape <- sf::read_sf("DC_Health_Planning_Neighborhoods") %>%
plot() 
```


```{r}
st_covered_by()
```

```{r download_data}
df1 <- read_csv("202109-capitalbikeshare-tripdata.csv") %>% clean_names()
```

Here We want to get the data filtered to just Sep 4
```{r manipulate_data}
df1a = df1 %>% 
  select(rideable_type, start = started_at, end = ended_at, start_station_name:end_lng) %>%
  pivot_longer(cols =  start:end,  names_to = "type", values_to = "time")%>%
  arrange(time) %>%
  mutate(inc = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  ), riders = cumsum(inc)) %>%
  mutate(day_of_year = yday(time)) %>%
  mutate(day_since_sep1 = day_of_year-243) %>%
  filter(day_since_sep1 == 4)
```

Here is the visualization
```{r visualization}
ggplot(df1a, mapping = aes(x = start_lat, y = start_lng)) +
  geom_point() #+ geom_sf(aes(neighborhoods_shape))
ggplot(df1a, mapping = aes(x = end_lat, y = end_lng)) +
  geom_point()
```
=======
---
title: "dc_covid"
author: "Jackie Budka"
date: "11/29/2021"
output: html_notebook
---



```{r load packages}
library(tidyverse)
library(tidymodels)
library(dsbox)
library(httr)
library(jsonlite)
library(lubridate)
library(sf)
library(tmap)
library(ggplot2)
```

```{r}
# url <- "https://opendata.arcgis.com"
# path <- "datasets/62f6f28dbfc74ae489fa3dcda9e94744_28.geojson"
# raw_data <- GET(url,path)
```
```{r}
neighborhoods_shape <- sf::read_sf("DC_Health_Planning_Neighborhoods") %>%
plot() 
```


```{r}
st_covered_by()
```

```{r download_data}
df1 <- read_csv("202109-capitalbikeshare-tripdata.csv") %>% clean_names()
```

Here We want to get the data filtered to just Sep 4
```{r manipulate_data}
df1a = df1 %>% 
  select(rideable_type, start = started_at, end = ended_at, start_station_name:end_lng) %>%
  pivot_longer(cols =  start:end,  names_to = "type", values_to = "time")%>%
  arrange(time) %>%
  mutate(inc = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  ), riders = cumsum(inc)) %>%
  mutate(day_of_year = yday(time)) %>%
  mutate(day_since_sep1 = day_of_year-243) %>%
  filter(day_since_sep1 == 4)
```

Here is the visualization
```{r visualization}
ggplot(df1a, mapping = aes(x = start_lat, y = start_lng)) +
  geom_point() #+ geom_sf(aes(neighborhoods_shape))
ggplot(df1a, mapping = aes(x = end_lat, y = end_lng)) +
  geom_point()
```
>>>>>>> c4a99e3966e28024f12b4b61717ea59d40702c7e
